---
output:
  md_document:
    variant: markdown_github
---

# Purpose

This is the root README for my Data Science practical, student number 24991422. It collates the four questions, each of which lives in its own folder and works on a standalone basis. The point of this file is to source each question's functions from its `code/` folder and render the key figures and tables here, so the whole submission can be read and checked from one place.

The workflow is the same throughout. Every operation is written as a function, one per file, kept in that question's `code/` folder and sourced together at the top of each section. The data sits in each question's `data/` folder and is not tracked by the repository, so the chunks below expect the files to be in place before the README knits.

### How to get started

I set the project up with the following, which creates the repository and the Texevier templates for the three written questions. Question 1 is an `officer` PowerPoint deck rather than a Texevier document, so it has no template.

```{r, eval = FALSE}
CHOSEN_LOCATION <- "~/DataScience/Practical/"
fmxdat::make_project(FilePath = glue::glue("{CHOSEN_LOCATION}Solution/"), ProjNam = "24991422")

Texevier::create_template(directory = glue::glue("{CHOSEN_LOCATION}Solution/24991422/"), template_name = "Question2")
Texevier::create_template(directory = glue::glue("{CHOSEN_LOCATION}Solution/24991422/"), template_name = "Question3")
Texevier::create_template(directory = glue::glue("{CHOSEN_LOCATION}Solution/24991422/"), template_name = "Question4")
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.align = "center", fig.width = 7, fig.height = 4,
                      dpi = 200, dev = "ragg_png", out.width = "100%")

rm(list = ls()); gc()    # clean the environment
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, lubridate, scales, ggrepel, knitr, kableExtra)
```

# Question 1, Coffee

A recommendation deck for stocking the Neelsie Coffee Hub, built in R with `officer` so the slides regenerate from the data. The deck itself is `Question1/Coffee_Hub_Recommendation.pptx`, and the charts inside it come from the functions sourced below. I reproduce them here from the same prepared data the deck uses.

My approach is to let the supplied reviews drive every claim so that nothing is hand-picked. `load_coffee` reads and cleans the file, `region_from_origin` rolls scattered growing regions up to countries, and `derive_keywords` ranks review words by the rating lift they carry rather than by a list I chose, after which `score_keywords` tags each coffee on those words. The four plots then turn that one prepared table into the roast, region-value, flavour and supplier views the deck recommends from, so each slide traces back to a number in the data.

```{r q1-prep}
list.files('Question1/code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

Coffee <- load_coffee("Question1/data/Coffee/Coffee.csv") %>% # load and fix the strange characters
    filter(Cost_Per_100g > 0) %>% # drop the odd zero-cost rows
    mutate(desc_all = replace_na(desc_all, "")) # a couple of reviews are missing

# roll growing regions up to countries
region_lookup <- c("Ethiopia"="Ethiopia","Kenya"="Kenya","Colombia"="Colombia","Panama"="Panama", "Hawai"="Hawaii","Guatemala"="Guatemala","Sumatra|Lintong|Sulawesi|Indonesia"="Indonesia", "Costa Rica"="Costa Rica","Brazil"="Brazil","Honduras"="Honduras","Rwanda"="Rwanda")

Coffee <- region_from_origin(Coffee, region_lookup) # tag each coffee's region
keywords <- derive_keywords(Coffee, n = 12) # indicator words, straight from the reviews
Coffee <- score_keywords(Coffee, keywords) # score each coffee on those words
```

Grouping every coffee by roast level and averaging the rating sets the roast strategy. The lighter roasts come out on top, with light and medium-light scoring highest and the rating sliding as the roast darkens, so the range should be built on lighter roasts.

```{r q1-roast}
plot_roast(Coffee)
```

Each growing country sits here by its average rating against its average price per 100g. East African origins like Kenya and Ethiopia land within a point of Panama while costing a fraction as much, and that gap is the core sourcing argument.

```{r q1-region}
plot_region_value(Coffee)
```

The fingerprint weighs how often each data-derived flavour word turns up in the best coffees against the shelf as a whole. Words like juicy and saturated are far more common among top-rated coffees, which marks them out as the notes worth stocking towards.

```{r q1-flavour}
plot_flavour_fingerprint(Coffee, keywords)
```

Ranking suppliers by their average rating pulls the affordable apart from the premium. A handful of roasters deliver high quality at low cost while a separate group holds the top of the price range, so the shelf can carry both a value tier and a showpiece tier.

```{r q1-supplier}
plot_supplier_leaderboard(Coffee)
```

# Question 2, Baby Names

How American culture writes itself into a cradle, read from the full state record of births from 1910 to 2014. The functions collate the national counts, measure how long the top names persist, then test music and screen against the surges directly using the Billboard and HBO datasets.

My approach splits the question into persistence and surges. `collate_names` and `national_totals` build the national series, `name_persistence` and `persistence_table` measure how firmly each year's top names hold their rank over the following years, and `find_spikes` pulls the sharp year-on-year jumps straight from the data. To test cause rather than assert it, `billboard_name_spikes` and `hbo_name_spikes` match those jumps against the Billboard and HBO datasets, `combine_matches` and `tag_spike_source` label each surge by what drove it, and `plot_event_study` pools the matches to read the average naming response around an event.

```{r q2-prep, cache = TRUE}
list.files('Question2/code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

nat <- collate_names("Question2/data/US_Baby_names/Baby_Names_By_US_State.rds") # national name counts

tot <- national_totals(nat) # collapsed across gender

charts <- read_rds("Question2/data/US_Baby_names/charts.rds") # weekly Billboard Hot 100

titles <- read_rds("Question2/data/US_Baby_names/HBO_titles.rds") # HBO titles and scores

credits <- read_rds("Question2/data/US_Baby_names/HBO_credits.rds") # HBO cast and characters

pers <- bind_rows(name_persistence(nat, "M"), name_persistence(nat, "F")) # persistence series

ptab <- persistence_table(pers) # summary table

spikes <- find_spikes(nat) # data-driven year-on-year surges

bb <- billboard_name_spikes(nat, charts) # singer-driven names

hbo <- hbo_name_spikes(nat, titles, credits) %>% # character-driven names
    mutate(across(where(is.character), ~ iconv(., to = "UTF-8", sub = ""))) # drop invalid bytes

matches <- combine_matches(bb, hbo) # names driven by both

sourced <- tag_spike_source(spikes, bb, hbo) # surges labelled by cause

fade_names <- spikes %>% 
    filter(Count >= 2000) %>% 
    slice_max(Growth, n = 6) %>% 
    pull(Name)

dist_names <- tot %>% 
    group_by(Name) %>%
    summarise(peak = max(Count), last = Count[which.max(Year)], .groups = "drop") %>%
    filter(peak >= 10000, last < 0.12 * peak) %>% 
    slice_max(peak, n = 6) %>%
    pull(Name)

bb_events  <- bb  %>% 
    slice_max(Ratio, n = 5) %>% 
    transmute(Name, Event_Year = Chart_Year)

hbo_events <- hbo %>% 
    slice_max(Ratio, n = 5) %>% 
    transmute(Name, Event_Year = release_year)
ev_events  <- matches %>% transmute(Name, Event_Year, Source)
```

For every year, the 25 most popular names are followed one, two and three years on, and the rank correlation is drawn by gender. The top names hold their position firmly through most of the century before loosening after 1990, and that loosening runs sharper for boys than for girls.

```{r q2-persist, fig.height = 3}
plot_persistence(pers)
```

```{r q2-ptab}
ptab %>% 
    kable(caption = "Mean rank persistence by gender and horizon, before and from 1990",
               digits = 3, col.names = c("Gender", "Horizon", "Pre-1990", "1990 onward", "Change"))
```

Here are the ten sharpest year-on-year surges, each tagged with the cause the datasets can attribute. The biggest movements are sudden spikes rather than slow drifts, and most of them line up with a song or a screen character.

```{r q2-surges}
sourced %>%
    slice_max(Growth, n = 10) %>%
    transmute(Year, Name, Before = Prev, After = Count, `Times bigger` = round(Growth, 1), Cause = Source) %>%
    kable(caption = "The ten sharpest year-on-year name surges, with the cause our datasets can attribute")
```

Every culturally driven name is placed by the year it surged, sized by the babies it reached at the peak and coloured by whether a singer or a character drove it. The surges cluster in the music and television era, and both sources throw up names that climb into the thousands.

```{r q2-bubble, fig.height = 4, fig.width = 7.5}
plot_spike_bubble(matches)
```

Following the sharpest surges across their whole life, rather than freezing them at the peak, tells a blunter story. Almost all of them are short-lived fashions that climb fast and fade inside a decade.

```{r q2-fade, fig.height = 3}
plot_trajectories(tot, fade_names, title = "Fade or stick", subtitle = "The sharpest surges are usually short-lived fashions")
```

This plot rescales the biggest names of the century to their own peak so the full arc is visible. It shows that even the largest names eventually fall to a fraction of their height, so size at the top is no guarantee of staying power.

```{r q2-dist, fig.height = 3.4}
plot_name_distribution(tot, dist_names)
```

This is the roll of singers whose first name jumped in newborns once they first cracked the Billboard top ten. The chart all but manufactures names here, with Sade and Rihanna climbing from almost nothing into the thousands.

```{r q2-bbtab}
bb %>% 
    slice_max(Ratio, n = 8) %>%
    transmute(Name, Artist = artist, `First charted` = Chart_Year, Before = Pre, Peak = Post, `Times bigger` = Ratio) %>%
    kable(caption = "Singers whose first name surged in babies after they first reached the Billboard top ten")
```

Plotting those same names from 1975 with a marker on each singer's first chart year makes the timing visible. It shows the babies arriving right after the music lands rather than before.

```{r q2-bbfig, fig.height = 2.9}
plot_trajectories(tot, bb_events$Name, events = bb_events, from = 1975,
                  title = "Named after the charts", subtitle = "Babies named for chart-topping singers")
```

This table lists the screen characters whose first name surged in babies after the title appeared. It shows the same effect from television and film, with names like Meadow after The Sopranos and Imani after Coming to America arriving within a year or two of release.

```{r q2-hbotab}
hbo %>% slice_max(Ratio, n = 8) %>%
    transmute(Name, Title = title, Released = release_year, Score = tmdb_score, Peak = Post, `Times bigger` = Ratio) %>%
    kable(caption = "Screen characters whose first name surged in babies after the title appeared")
```

Plotting the character names from 1980 with a marker on each title's release does the same for screen. It shows the same short lag, with the names rising within a year or two of the title.

```{r q2-hbofig, fig.height = 2.9}
plot_trajectories(tot, hbo_events$Name, events = hbo_events, from = 1980,
                  title = "Named after the screen", subtitle = "Babies named for popular screen characters")
```

Pooling every matched name on the year its trigger landed and averaging around it gives the event study that closes the question. Naming sits flat before the trigger and jumps in the years just after, the cleanest single piece of evidence that culture moves names.

```{r q2-event, fig.height = 2.9}
plot_event_study(tot, ev_events)
```

# Question 3, Loans and Credit

What drives default across a million anonymised Lending Club loans. Default is defined as a loan charged off against one fully paid, and the analysis stays descriptive, reading default rates across the factors the Institute cares about. The data is large, so this chunk is cached.

My approach starts with what default means, so I resolve each loan to charged off against fully paid and drop the unresolved cases, which `load_loans` and `prep_default` handle. From there the work stays descriptive rather than modelled, reading the default rate across the factors the Institute named, with `plot_grade`, `plot_dti` and `plot_states` covering the lender's own grade, the debt-to-income lever and the geographic claim about Texas. The aim is to test the agency's beliefs against the data rather than to build a classifier.

```{r q3-prep, cache = TRUE}
list.files('Question3/code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

loans <- load_loans("Question3/data/Loan_Cred/loan_data.rds") %>% # the one-million-row extract
    prep_default()# resolved loans with the default flag

sp <- function(tab) round(diff(range(tab$Default)), 1) # pp spread of a summary
bandit <- function(x, b, l) cut(x, b, labels = l) # quick banding helper

drivers <- tibble(# one row per factor, ranked by its spread
  Driver = c("Credit grade","Interest rate","Loan term","Debt-to-income",
             "Recent inquiries","Revolving use","Home ownership","Employment length","Public records"),
  Spread = c( # for each factor, the gap in default rate between its safest and riskiest band
    sp(default_rate(loans, grade)),  # grade A through G, grouped directly
    sp(default_rate(mutate(loans,# interest rate has to be banded into a category first
                           b = bandit(int_rate, c(0,10,15,20,100), c("<10","10-15","15-20","20+"))), b)),  # cut int_rate into four bands, then default rate per band
    sp(default_rate(loans, term)), # 36 vs 60 month term, grouped directly
    sp(default_rate(filter(loans, # drop the loans with no usable dti first
                           !is.na(dti)) %>%
                        mutate(b = bandit(dti, c(0,10,20,30,60), c("0-10","10-20","20-30","30+"))), b)), # band dti, then default rate per band
    sp(default_rate(mutate(loans, b = bandit(inq_last_6mths, c(-1,0,1,2,99), c("0","1","2","3+"))), b)),  # recent credit inquiries banded 0/1/2/3+
    sp(default_rate(mutate(loans, b = bandit(revol_util, c(-1,25,50,75,1000), c("<25","25-50","50-75","75+"))), b)), # revolving-credit use banded into quarters
    sp(default_rate(loans, home_ownership)), # own / mortgage / rent, grouped directly
    sp(default_rate(loans, emp_length)), # employment-length category, grouped directly
    sp(default_rate(mutate(loans, b = bandit(pub_rec, c(-1,0,99), c("0","1+"))), b)))) %>% # public records collapsed to none vs any
  arrange(desc(Spread)) # biggest spread first, so the strongest driver sits at the top

natg  <- loans %>% 
    group_by(grade) %>% 
    summarise(g = mean(default), .groups = "drop")

state <- loans %>% 
    left_join(natg, by = "grade") %>% 
    group_by(addr_state) %>% # observed vs grade-mix expectation
    summarise(Observed = round(mean(default) * 100, 1), Expected = round(mean(g) * 100, 1),
              Loans = n(), .groups = "drop") %>%
    filter(Loans >= 500) %>% 
    mutate(Excess = round(Observed - Expected, 1))

dti_band <- loans %>% filter(!is.na(dti)) %>% # default rate by debt-to-income band
    mutate(band = cut(dti, c(0,10,15,20,25,30,35,40,60), right = FALSE,
                      labels = c("0-10","10-15","15-20","20-25","25-30","30-35","35-40","40+"))) %>%
    group_by(band) %>% 
    summarise(Default = round(mean(default) * 100, 1), .groups = "drop")

cap_for  <- function(tol){ ok <- dti_band$band[dti_band$Default <= tol]
    edges <- c("0-10"=10,"10-15"=15,"15-20"=20,"20-25"=25,"25-30"=30,"30-35"=35,"35-40"=40,"40+"=60)
    if(length(ok)) unname(edges[as.character(tail(ok, 1))]) else NA } # upper edge of the highest safe band
cap_tab  <- tibble(`Default tolerance` = c("18%","20%","22%","25%"),
                   `DTI cap` = c(cap_for(18), cap_for(20), cap_for(22), cap_for(25)))
```

Default is read off against the grade the lender already assigns, from A through G. It climbs without a single reversal, from about seven percent at grade A to nearly sixty percent at grade G, a spread that dwarfs every other factor in the data.

```{r q3-grade}
plot_grade(loans)
```

Ranking every factor by the gap between its safest and riskiest band puts that dominance in one place. Grade tops the list by a wide margin, and the things the Institute worried about, ownership and employment among them, sit far down it.

```{r q3-drivers}
drivers %>% 
    kable(caption = "Risk drivers ranked by the gap in default rate between their safest and riskiest band")
```

The next view reads the default rate across bands of the debt-to-income ratio. Default rises smoothly as the ratio grows rather than breaking at any threshold, so where to draw the cap is a tolerance call rather than a natural line.

```{r q3-dti}
plot_dti(loans)
```

Turning that curve into a decision, this reads off the highest debt-to-income band that still clears each default tolerance. A stricter ceiling buys a lower cap, and no single value stands out as the obvious place to cut.

```{r q3-cap}
cap_tab %>% kable(caption = "The highest debt-to-income band that still clears each default tolerance")
```

Each state's actual default rate is set beside what its grade mix alone would predict. Most of the gap between states turns out to be grade composition, and Texas in particular lands almost exactly on the national average rather than standing apart.

```{r q3-states, fig.height = 6}
plot_states(loans)
```

Listing the states with the largest excess default in either direction, with Texas set alongside, separates real geography from grade mix. The extremes are modest once grade is accounted for, and Texas carries almost no excess at all.

```{r q3-state-tab}
bind_rows(slice_max(state, Excess, n = 4), slice_min(state, Excess, n = 4), filter(state, addr_state == "TX")) %>%
    distinct() %>% arrange(desc(Excess)) %>%
    kable(col.names = c("State", "Observed %", "Expected %", "Loans", "Excess pp"),
          caption = "States with the largest positive and negative excess default, with Texas for reference")
```

# Question 4, Netflix

A read on the Netflix catalogue for a team planning its own streaming service. The functions parse the genres and countries, then map where the films come from, what each country makes, how they rate and how long they run, with HBO as a benchmark.

My approach parses the data before reading anything from it, so `load_streaming` pulls genre and country out of the raw text fields and `prep_movies` keeps films up to 2022. The plots then read the catalogue from a few angles, where titles come from, what each country specialises in, how acclaim and popularity diverge and how long films run, and `plot_hbo` sets all of it against HBO as a benchmark. `distinctive_words` cross-checks the genre map against the language of the descriptions so the picture does not rest on the genre tags alone.

```{r q4-prep}
list.files('Question4/code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

titles <- load_streaming("Question4/data/netflix/titles.rds") # IMDb-sourced titles

movies <- prep_movies(titles) # films up to 2022

hbo <- load_streaming("Question4/data/HBO_titles.rds") %>% 
    filter(type == "MOVIE") # HBO films

topc <- movies %>% 
    unnest(country) %>% 
    count(country, sort = TRUE) %>% 
    slice_head(n = 8) %>% 
    pull(country)

topg <- movies %>% 
    unnest(genre) %>% 
    count(genre, sort = TRUE) %>%
    slice_head(n = 8) %>% 
    pull(genre)

content_g <- c("drama","comedy","thriller","romance","action","documentation","crime")
```

Ranking the production countries by how many films each puts into the catalogue maps its centre of gravity. The library is American at its core but far from American alone, with India a clear second.

```{r q4-countries}
plot_countries(movies, 10)
```

This plot gives the share of each country's films that falls in each major genre. It shows that every tradition specialises differently, with India leaning on drama and romance, Japan on action, and the United States and Britain on documentaries.

```{r q4-heat}
plot_genre_country(movies, topc, topg)
```

Each genre is placed by its typical critic score against its typical audience size. Acclaim and popularity pull apart here, with documentaries rating highly on small audiences and thrillers doing the reverse, and that split is the most useful signal for a new entrant.

```{r q4-ratings}
plot_ratings(movies, topg)
```

Running time, compared across production countries, turns out to carry a signature of its own. Indian films run far longer than the American median, a difference clear enough to read off the chart.

```{r q4-length}
plot_length(movies, topc)
```

Setting the typical Netflix score against HBO within each genre gives a like-for-like benchmark. Netflix trades quality for breadth, rating below HBO in every genre rather than only on the overall average.

```{r q4-hbo}
plot_hbo(movies, hbo, content_g)
```

This table lists the words most distinctive to each genre's descriptions, measured by how much more they appear there than elsewhere. It shows that each genre reaches for its own vocabulary, which confirms the genre map from the language up rather than from the tags alone.

```{r q4-words}
distinctive_words(movies, content_g) %>%
    group_by(genre) %>%
    summarise(`Distinctive words` = str_c(word, collapse = ", "), .groups = "drop") %>%
    kable()
```
