---
output:
  md_document:
    variant: markdown_github
---

# Purpose

This is the root README for my Data Science practical, student number 24991422. It collates the four questions, each of which lives in its own folder and works on a standalone basis. The point of this file is to source each question's functions from its `code/` folder and render the key figures and tables here, so the whole submission can be read and checked from one place.

The workflow is the same throughout. Every operation is written as a function, one per file, kept in that question's `code/` folder and sourced together at the top of each section. The data sits in each question's `data/` folder and is not tracked by the repository, so the chunks below expect the files to be in place before the README knits.

### How to get started

I set the project up with the following, which creates the repository and the Texevier templates for the three written questions. Question 1 is an `officer` PowerPoint deck rather than a Texevier document, so it has no template.

```{r, eval = FALSE}
CHOSEN_LOCATION <- "~/DataScience/Practical/"
fmxdat::make_project(FilePath = glue::glue("{CHOSEN_LOCATION}Solution/"), ProjNam = "24991422")

Texevier::create_template(directory = glue::glue("{CHOSEN_LOCATION}Solution/24991422/"), template_name = "Question2")
Texevier::create_template(directory = glue::glue("{CHOSEN_LOCATION}Solution/24991422/"), template_name = "Question3")
Texevier::create_template(directory = glue::glue("{CHOSEN_LOCATION}Solution/24991422/"), template_name = "Question4")
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.align = "center", fig.width = 7, fig.height = 4,
                      dpi = 200, dev = "ragg_png", out.width = "100%")

rm(list = ls()); gc()    # clean the environment
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, lubridate, scales, ggrepel, knitr, kableExtra)
```

# Question 1, Coffee

A recommendation deck for stocking the Neelsie Coffee Hub, built in R with `officer` so the slides regenerate from the data. The deck itself is `Question1/Coffee_Hub_Recommendation.pptx`, and the charts inside it come from the functions sourced below. I reproduce them here from the same prepared data the deck uses.

My approach is to let the supplied reviews drive every claim so that nothing is hand-picked. `load_coffee` reads and cleans the file, `region_from_origin` rolls scattered growing regions up to countries, and `derive_keywords` ranks review words by the rating lift they carry rather than by a list I chose, after which `score_keywords` tags each coffee on those words. The four plots then turn that one prepared table into the roast, region-value, flavour and supplier views the deck recommends from, so each slide traces back to a number in the data.

```{r q1-prep}
list.files('Question1/code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

Coffee <- load_coffee("Question1/data/Coffee/Coffee.csv") %>% # load and fix the strange characters
    filter(Cost_Per_100g > 0) %>% # drop the odd zero-cost rows
    mutate(desc_all = replace_na(desc_all, "")) # a couple of reviews are missing

# roll growing regions up to countries
region_lookup <- c("Ethiopia"="Ethiopia","Kenya"="Kenya","Colombia"="Colombia","Panama"="Panama", "Hawai"="Hawaii","Guatemala"="Guatemala","Sumatra|Lintong|Sulawesi|Indonesia"="Indonesia", "Costa Rica"="Costa Rica","Brazil"="Brazil","Honduras"="Honduras","Rwanda"="Rwanda")

Coffee <- region_from_origin(Coffee, region_lookup) # tag each coffee's region
keywords <- derive_keywords(Coffee, n = 12) # indicator words, straight from the reviews
Coffee <- score_keywords(Coffee, keywords) # score each coffee on those words
```

This plot gives the average rating of every coffee grouped by roast level. It shows that the lighter roasts come out on top, with light and medium-light rating highest and the score sliding as the roast darkens, so the range should be anchored on lighter roasts.

```{r q1-roast}
plot_roast(Coffee)
```

This plot places each growing country by its average rating against its average price per 100g. It shows that East African origins like Kenya and Ethiopia rate within a point of Panama while costing a fraction as much, which is the core sourcing argument.

```{r q1-region}
plot_region_value(Coffee)
```

This plot compares how often each data-derived flavour word appears in the very best coffees against the shelf as a whole. It shows that words like juicy and saturated are far more common among top-rated coffees, so those are the notes worth stocking towards.

```{r q1-flavour}
plot_flavour_fingerprint(Coffee, keywords)
```

This plot ranks suppliers by their average rating and separates the affordable from the premium. It shows that a few roasters deliver high quality at low cost while a different group sits at the top of the price range, so the shelf can carry both a value tier and a showpiece tier.

```{r q1-supplier}
plot_supplier_leaderboard(Coffee)
```

# Question 2, Baby Names

How American culture writes itself into a cradle, read from the full state record of births from 1910 to 2014. The functions collate the national counts, measure how long the top names persist, then test music and screen against the surges directly using the Billboard and HBO datasets.

My approach splits the question into persistence and surges. `collate_names` and `national_totals` build the national series, `name_persistence` and `persistence_table` measure how firmly each year's top names hold their rank over the following years, and `find_spikes` pulls the sharp year-on-year jumps straight from the data. To test cause rather than assert it, `billboard_name_spikes` and `hbo_name_spikes` match those jumps against the Billboard and HBO datasets, `combine_matches` and `tag_spike_source` label each surge by what drove it, and `plot_event_study` pools the matches to read the average naming response around an event.

```{r q2-prep, cache = TRUE}
list.files('Question2/code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

nat <- collate_names("Question2/data/US_Baby_names/Baby_Names_By_US_State.rds") # national name counts

tot <- national_totals(nat) # collapsed across gender

charts <- read_rds("Question2/data/US_Baby_names/charts.rds") # weekly Billboard Hot 100

titles <- read_rds("Question2/data/US_Baby_names/HBO_titles.rds") # HBO titles and scores

credits <- read_rds("Question2/data/US_Baby_names/HBO_credits.rds") # HBO cast and characters

pers <- bind_rows(name_persistence(nat, "M"), name_persistence(nat, "F")) # persistence series

ptab <- persistence_table(pers) # summary table

spikes <- find_spikes(nat) # data-driven year-on-year surges

bb <- billboard_name_spikes(nat, charts) # singer-driven names

hbo <- hbo_name_spikes(nat, titles, credits) %>% # character-driven names
    mutate(across(where(is.character), ~ iconv(., to = "UTF-8", sub = ""))) # drop invalid bytes

matches <- combine_matches(bb, hbo) # names driven by both

sourced <- tag_spike_source(spikes, bb, hbo) # surges labelled by cause

fade_names <- spikes %>% 
    filter(Count >= 2000) %>% 
    slice_max(Growth, n = 6) %>% 
    pull(Name)

dist_names <- tot %>% 
    group_by(Name) %>%
    summarise(peak = max(Count), last = Count[which.max(Year)], .groups = "drop") %>%
    filter(peak >= 10000, last < 0.12 * peak) %>% 
    slice_max(peak, n = 6) %>%
    pull(Name)

bb_events  <- bb  %>% 
    slice_max(Ratio, n = 5) %>% 
    transmute(Name, Event_Year = Chart_Year)

hbo_events <- hbo %>% 
    slice_max(Ratio, n = 5) %>% 
    transmute(Name, Event_Year = release_year)
ev_events  <- matches %>% transmute(Name, Event_Year, Source)
```

This plot tracks the rank correlation of each year's 25 most popular names one, two and three years later, split by gender. It shows that the top names hold their position firmly through most of the century and then loosen after 1990, more sharply for boys than for girls.

```{r q2-persist, fig.height = 3}
plot_persistence(pers)
```

```{r q2-ptab}
ptab %>% 
    kable(caption = "Mean rank persistence by gender and horizon, before and from 1990",
               digits = 3, col.names = c("Gender", "Horizon", "Pre-1990", "1990 onward", "Change"))
```

This table lists the ten sharpest year-on-year name surges with the cause our datasets can attribute to each. It shows that the biggest movements are sudden spikes rather than slow drifts, and that most of them line up with a song or a screen character.

```{r q2-surges}
sourced %>%
    slice_max(Growth, n = 10) %>%
    transmute(Year, Name, Before = Prev, After = Count, `Times bigger` = round(Growth, 1), Cause = Source) %>%
    kable(caption = "The ten sharpest year-on-year name surges, with the cause our datasets can attribute")
```

This plot places each culturally driven name by the year it surged, sized by how many babies took it at the peak and coloured by whether a singer or a screen character drove it. It shows that the surges cluster in the music and television era and that both sources produce names reaching into the thousands.

```{r q2-bubble, fig.height = 4, fig.width = 7.5}
plot_spike_bubble(matches)
```

This plot follows the names with the sharpest surges across their whole life rather than at the peak alone. It shows that almost all of them are short-lived fashions that climb fast and fade within about a decade.

```{r q2-fade, fig.height = 3}
plot_trajectories(tot, fade_names, title = "Fade or stick", subtitle = "The sharpest surges are usually short-lived fashions")
```

This plot rescales the biggest names of the century to their own peak so the full arc is visible. It shows that even the largest names eventually fall to a fraction of their height, so size at the top is no guarantee of staying power.

```{r q2-dist, fig.height = 3.4}
plot_name_distribution(tot, dist_names)
```

This table lists the singers whose first name surged in babies after they first reached the Billboard top ten. It shows the chart almost creating names, with Sade and Rihanna climbing from near nothing into the thousands.

```{r q2-bbtab}
bb %>% 
    slice_max(Ratio, n = 8) %>%
    transmute(Name, Artist = artist, `First charted` = Chart_Year, Before = Pre, Peak = Post, `Times bigger` = Ratio) %>%
    kable(caption = "Singers whose first name surged in babies after they first reached the Billboard top ten")
```

Plotting those same names from 1975 with a marker on each singer's first chart year makes the timing visible. It shows the babies arriving right after the music lands rather than before.

```{r q2-bbfig, fig.height = 2.9}
plot_trajectories(tot, bb_events$Name, events = bb_events, from = 1975,
                  title = "Named after the charts", subtitle = "Babies named for chart-topping singers")
```

This table lists the screen characters whose first name surged in babies after the title appeared. It shows the same effect from television and film, with names like Meadow after The Sopranos and Imani after Coming to America arriving within a year or two of release.

```{r q2-hbotab}
hbo %>% slice_max(Ratio, n = 8) %>%
    transmute(Name, Title = title, Released = release_year, Score = tmdb_score, Peak = Post, `Times bigger` = Ratio) %>%
    kable(caption = "Screen characters whose first name surged in babies after the title appeared")
```

Plotting the character names from 1980 with a marker on each title's release does the same for screen. It shows the same short lag, with the names rising within a year or two of the title.

```{r q2-hbofig, fig.height = 2.9}
plot_trajectories(tot, hbo_events$Name, events = hbo_events, from = 1980,
                  title = "Named after the screen", subtitle = "Babies named for popular screen characters")
```

This plot pools every matched name on the year its trigger landed and averages the response around it. It shows a clean event-study shape, with naming flat before the trigger and jumping in the years just after, which is the closing evidence that culture moves names.

```{r q2-event, fig.height = 2.9}
plot_event_study(tot, ev_events)
```

# Question 3, Loans and Credit

What drives default across a million anonymised Lending Club loans. Default is defined as a loan charged off against one fully paid, and the analysis stays descriptive, reading default rates across the factors the Institute cares about. The data is large, so this chunk is cached.

My approach starts with what default means, so I resolve each loan to charged off against fully paid and drop the unresolved cases, which `load_loans` and `prep_default` handle. From there the work stays descriptive rather than modelled, reading the default rate across the factors the Institute named, with `plot_grade`, `plot_dti` and `plot_states` covering the lender's own grade, the debt-to-income lever and the geographic claim about Texas. The aim is to test the agency's beliefs against the data rather than to build a classifier.

```{r q3-prep, cache = TRUE}
list.files('Question3/code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

loans <- load_loans("Question3/data/Loan_Cred/loan_data.rds") %>% # the one-million-row extract
    prep_default() # resolved loans with the default flag
```

This plot gives the default rate for each loan grade the lender assigns, from A to G. It shows that default climbs without a single reversal from roughly seven percent at grade A to nearly sixty percent at grade G, a spread that dwarfs every other factor in the data.

```{r q3-grade}
plot_grade(loans)
```

This plot gives the default rate across bands of the debt-to-income ratio. It shows that default rises smoothly as the ratio grows rather than jumping at any threshold, so where to cap it is a tolerance choice rather than a natural break.

```{r q3-dti}
plot_dti(loans)
```

This plot sets each state's default rate next to what its grade mix alone would predict. It shows that most of the gap between states is just grade composition, and that Texas sits almost exactly on the national average rather than apart from it.

```{r q3-states, fig.height = 6}
plot_states(loans)
```

# Question 4, Netflix

A read on the Netflix catalogue for a team planning its own streaming service. The functions parse the genres and countries, then map where the films come from, what each country makes, how they rate and how long they run, with HBO as a benchmark.

My approach parses the data before reading anything from it, so `load_streaming` pulls genre and country out of the raw text fields and `prep_movies` keeps films up to 2022. The plots then read the catalogue from a few angles, where titles come from, what each country specialises in, how acclaim and popularity diverge and how long films run, and `plot_hbo` sets all of it against HBO as a benchmark. `distinctive_words` cross-checks the genre map against the language of the descriptions so the picture does not rest on the genre tags alone.

```{r q4-prep}
list.files('Question4/code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

titles <- load_streaming("Question4/data/netflix/titles.rds") # IMDb-sourced titles

movies <- prep_movies(titles) # films up to 2022

hbo <- load_streaming("Question4/data/HBO_titles.rds") %>% 
    filter(type == "MOVIE") # HBO films

topc <- movies %>% 
    unnest(country) %>% 
    count(country, sort = TRUE) %>% 
    slice_head(n = 8) %>% 
    pull(country)

topg <- movies %>% 
    unnest(genre) %>% 
    count(genre, sort = TRUE) %>%
    slice_head(n = 8) %>% 
    pull(genre)

content_g <- c("drama","comedy","thriller","romance","action","documentation","crime")
```

This plot ranks the production countries by how many films each contributes to the catalogue. It shows that the library is American at its core but far from American alone, with India a clear second.

```{r q4-countries}
plot_countries(movies, 10)
```

This plot gives the share of each country's films that falls in each major genre. It shows that every tradition specialises differently, with India leaning on drama and romance, Japan on action, and the United States and Britain on documentaries.

```{r q4-heat}
plot_genre_country(movies, topc, topg)
```

This plot places each genre by its typical critic score against its typical audience size. It shows that acclaim and popularity pull apart, with documentaries scoring high on small audiences and thrillers doing the reverse, which is the most useful signal for a new entrant.

```{r q4-ratings}
plot_ratings(movies, topg)
```

This plot compares the typical running time of films by production country. It shows that length is a national fingerprint too, with Indian films running far longer than the American median.

```{r q4-length}
plot_length(movies, topc)
```

This plot sets the typical Netflix score against HBO within each genre. It shows that Netflix trades quality for breadth, rating below HBO in every genre rather than only on average.

```{r q4-hbo}
plot_hbo(movies, hbo, content_g)
```

This table lists the words most distinctive to each genre's descriptions, measured by how much more they appear there than elsewhere. It shows that each genre reaches for its own vocabulary, which confirms the genre map from the language up rather than from the tags alone.

```{r q4-words}
distinctive_words(movies, content_g) %>%
    group_by(genre) %>%
    summarise(`Distinctive words` = str_c(word, collapse = ", "), .groups = "drop") %>%
    kable()
```
